---
title: Testing some MDS solutions on the data 
output: html_document
---


```{r}
suppressMessages(library(tidyverse))

rock_images <- "120 Rock Images/*.png"
judgements <- "Similarity Judgments Data"
mds_file <- "MDS/mds_120_supplemental_dims.txt"
```

## Original MDS coordinates

```{r}
# original MDS coordinates for 120 Rocks Set
mds_obs <- suppressMessages(read_tsv(mds_file, col_names=FALSE, skip=1))

cat("Number of rocks:     ", nrow(mds_obs))
cat("\nNumber of dimensions:", ncol(mds_obs))
```

## Replicated MDS coordinates with R smacof package

```{r}
# replicated MDS coordinates for 120 Rocks Set

# load judgements into 1 dataframe
suppressMessages(df_jm <- list.files(path = paste(getwd(),"/", judgements, sep=""), pattern = 'rocks_similarity_120_*', full.names = T) %>%
  map_dfr(read_table))

df_jm <- df_jm %>%
 mutate(
    Rock1Type = sprintf("%02d", as.numeric(Rock1Type)),
    Rock2Type = sprintf("%02d", as.numeric(Rock2Type)),
    Rock1Token = sprintf("%02d", as.numeric(Rock1Token)),
    Rock2Token = sprintf("%02d", as.numeric(Rock2Token))
 ) %>%
 mutate(
  Rock1 = paste(Rock1Type, Rock1Token, sep="_"),
  Rock2 = paste(Rock2Type, Rock2Token, sep="_")
 ) %>%
 select(Rock1, Rock2, Rating)

df_jm$Rating <- as.numeric(df_jm$Rating)
```


```{r}
suppressMessages(library(smacof))
```

```{r}
# Correct order of Rocks so that Rock1 has always a lower number. 
df_jm$reverse <- FALSE
df_jm[df_jm$Rock1 > df_jm$Rock2, "reverse"] <- TRUE
reversed <- df_jm[df_jm$reverse, c("Rock1", "Rock2")]
df_jm[df_jm$reverse, "Rock1"] <- reversed$Rock2
df_jm[df_jm$reverse, "Rock2"] <- reversed$Rock1
df_jm <- select(df_jm, -reverse)

# sort
df_jm <- arrange(df_jm, Rock1, Rock2)

df_jm_agg <- df_jm %>%
    group_by(Rock1, Rock2) %>%
    summarise(Rating = mean(Rating))

rocks <- unique(c(unique(df_jm_agg$Rock1), unique(df_jm_agg$Rock2)))

# empty matrix
m_sim <- matrix(nrow=length(rocks), ncol=length(rocks))

# fill the matrix
for (m in 1:length(rocks)) {
  for (n in 1:length(rocks)) {
      if (rocks[m]>rocks[n]) {
          rating_value <- pull(df_jm_agg[(df_jm_agg$Rock1==rocks[n]) & (df_jm_agg$Rock2==rocks[m]), "Rating"])
      } else {
          rating_value <- pull(df_jm_agg[(df_jm_agg$Rock1==rocks[m]) & (df_jm_agg$Rock2==rocks[n]), "Rating"])
      }
      if (length(rating_value) == 0) rating_value <- NA
      m_sim[m, n] <- rating_value
  } 
}
```


```{r}
m_dist <- sim2diss(m_sim, method = 10)
```


```{r}
set.seed(298)

# standard solution
res <- mds(m_dist, ndim=8, type="ordinal", principal=TRUE, init="random", itmax=2000)
mds_repl <- res$conf
mds_repl <- scales::rescale(mds_repl, to=c(-5, 5)) # rescale to original mds coords min max
write.table(mds_repl, "MDS/mds_120_replication_r.txt", row.names = FALSE) # write to file

# bootstrapped solution
# No difference visible between bootstrap and non-bootstrap method
# res_boot <- bootmds(res, m_sim, method.dat = "euclidean", nrep = 50)
```

```{r}
mds_obs <- read.table("MDS/mds_120_supplemental_dims.txt", skip=1, header=FALSE)
mds_obs <- select(mds_obs, -c(names(mds_obs)[9:13]))
```

### Try to find similar solution to original by combinating ...
... different seeds and dimension orders and correlate these solutions with the original mds coordinates

```{r}
perm <- function(v) {
  n <- length(v)
  if (n == 1) v
  else {
    X <- NULL
    for (i in 1:n) X <- rbind(X, cbind(v[i], perm(v[-i])))
    X
  }
}

dim_permutations <- perm(1:8)
```

```{r}
best_seed <- 115 # checked until seed 410 with other distance matrix
best_value <- 0.5697673
best_combination <- c(2,4,8,3,5,1,7,6)
```

```{r}
best_seed <- 0 # checked until seed 410
best_value <- 0
best_combination <- c(2,4,8,3,5,1,7,6)
```


```{r}
for (seed in 1:800) {
    
    set.seed(seed)
    res <- mds(m_dist, ndim=8, type="ordinal", principal=TRUE, init="random", itmax=5000, eps = 1e-09)
    mds_repl <- res$conf
    mds_repl <- scales::rescale(mds_repl, to=c(-5, 5))
    if (seed %% 5 == 0) {
        cat(seed, "...")
    }
    for (j in 1:nrow(dim_permutations)) {
        dim_repl <- dim_permutations[j, ]
        rs <- c()
        for (i in 1:8) {
            r <- cor(mds_obs[, i], mds_repl[, dim_repl[i]]) 
            rs <- append(rs, r)
        }

        if (mean(rs) > best_value) {
            best_seed <- seed 
            best_value <- mean(rs) 
            best_combination <- dim_repl
            cat("\n", best_seed,": ", best_value, ", Dim repl: ", best_combination, sep="")
        }
    }
}

cat("\n\nFinished.\n\n", best_seed,": ", best_value, ", Dim repl: ", best_combination, sep="")
```

#### Store mds with optimal random seed and swap dimensions

```{r}
set.seed(best_seed)
combination <- best_combination #c(2,4,8,3,5,1,6,7)

# run mds
res <- mds(m_dist, ndim=8, type="ordinal", principal=TRUE, init="random", itmax=5000, eps = 1e-09)
mds_repl <- res$conf
mds_repl <- scales::rescale(mds_repl, to=c(-5, 5)) # rescale to original mds coords min max

# swap dimensions
mds_repl <- mds_repl[,  combination]

# write to file
write.table(mds_repl, "MDS/mds_120_replication_r.txt", row.names = FALSE) # write to file
```

## Plot results

results are plotted in python together with the scikit-learn results in `mds_replication.ipynb`

